Declarative Pearl: Lightweight Probabilistic Programming in Curry

### 2015-09-15 Paper/paper.tex
Fix typo semantic/semantics

Contributions:
\begin{itemize}
\item We present \emph{PFLP} a lightweight library for probabilistic
  programming in the functional logic language Curry. %
  The library's core consists of ... data structures as well as
  ... functions to work on these structures, overall the library needs
  less than 30 lines of code. %
\item We can reduce the library code to such a minimum by reusing Curry's
  semantic. %
  The best part is: no further mechanisms are necessary to enable
  probabilistic computations in Curry. %
  That said, we show high connections between current probability
  programming languages and Curry, including lazy evaluation,
  nondeterministic values as well as corresponding built-in search
  abilities and call-time choice. %
\item We present examples that emphasise the benefits of Curry's
  representation of nondeterminism, which enables early pruning of the
  search space, compared to Haskell's MonadPlus instance for lists,
  which is often used to model nondeterminism. %
\item It is fair to assume that a wide range of tools existing for
  Curry can be applied to probabilistic languages as well. %
  For instance, \cite{Brassel-BIO} has already presented work on a
  debugger for Curry that works well with nondeterminism and free
  variables. %
\end{itemize}

@@

\begin{itemize}
\item PPL vs DSL\\
  Working within a host language enables us to focus on the
  problem itself and frees us from reimplementing standard primitives
  (IO, data structures), essential design decisions (syntax,
  parser). %
  % Based on which criteria do you choose the host language? %
  % As for probabilistic languages that started growing over time,
  % libraries that enable probabilistic programming exist in the whole
  % variety of paradigms: declarative as well as imperative. %
  % For a start, you obviously choose the paradigm that has the clearest
  % structure and enables straight-forward programming. %
  % We want to address a wide range of developers and even
  % non-PL-experts, thus, we choose a paradigm than can be learnt
  % easily. %
  % Thus, obviously, you choose a declarative programming language. %
  % By choosing Curry, we argue that it is sufficient to reuse built-in
  % features in order to cover the essence of probabilistic programs,
  % especially its semantic. %
\item describe underlying implementation: core data types and
  functions
\item query simple question for example from introduction
\item connect to Martin Erwig's PFP approach
\item Nondeterminism and \code{SearchTree}
\end{itemize}

@@

\begin{itemize}
\item SetFunctions (memory allocation; usage as ``first-class
  citizen'', i.e., check semantics)
\item ``outside'' nondeterminism
\item sampling technique for nondeterministic functions in Curry
\item reevaluate Currys semantic in constrast to Sato's distribution
  semantic
\item reactivate debugger for KiCS2/PAKCS
\end{itemize}

>>>>

Contributions:
\begin{itemize}
\item We present \emph{PFLP} a lightweight library for probabilistic
  programming in the functional logic language Curry. %
  The library's core consists of ... data structures as well as
  ... functions to work on these structures, overall the library needs
  less than 30 lines of code. %
\item We can reduce the library code to such a minimum by reusing Curry's
  semantics. %
  The best part is: no further mechanisms are necessary to enable
  probabilistic computations in Curry. %
  That said, we show high connections between current probability
  programming languages and Curry, including lazy evaluation,
  nondeterministic values as well as corresponding built-in search
  abilities and call-time choice. %
\item We present examples that emphasise the benefits of Curry's
  representation of nondeterminism, which enables early pruning of the
  search space, compared to Haskell's MonadPlus instance for lists,
  which is often used to model nondeterminism. %
\item It is fair to assume that a wide range of tools existing for
  Curry can be applied to probabilistic languages as well. %
  For instance, \cite{Brassel-BIO} has already presented work on a
  debugger for Curry that works well with nondeterminism and free
  variables. %
\end{itemize}

@@

\begin{itemize}
\item PPL vs DSL\\
  Working within a host language enables us to focus on the
  problem itself and frees us from reimplementing standard primitives
  (IO, data structures), essential design decisions (syntax,
  parser). %
  % Based on which criteria do you choose the host language? %
  % As for probabilistic languages that started growing over time,
  % libraries that enable probabilistic programming exist in the whole
  % variety of paradigms: declarative as well as imperative. %
  % For a start, you obviously choose the paradigm that has the clearest
  % structure and enables straight-forward programming. %
  % We want to address a wide range of developers and even
  % non-PL-experts, thus, we choose a paradigm than can be learnt
  % easily. %
  % Thus, obviously, you choose a declarative programming language. %
  % By choosing Curry, we argue that it is sufficient to reuse built-in
  % features in order to cover the essence of probabilistic programs,
  % especially its semantics. %
\item describe underlying implementation: core data types and
  functions
\item query simple question for example from introduction
\item connect to Martin Erwig's PFP approach
\item Nondeterminism and \code{SearchTree}
\end{itemize}

@@

\begin{itemize}
\item SetFunctions (memory allocation; usage as ``first-class
  citizen'', i.e., check semantics)
\item ``outside'' nondeterminism
\item sampling technique for nondeterministic functions in Curry
\item reevaluate Currys semantics in constrast to Sato's distribution
  semantics
\item reactivate debugger for KiCS2/PAKCS
\end{itemize}
